When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2017-11-07 paper/paper.tex
More big data

To show that a pure subgraph isomorphism approach is feasible, with no indexing or supporting
preprocessing, we look at four datasets commonly used in graph indexing evaluations. We do not claim
that these are high-quality datasets or that the associated queries are sensible, merely that
following the example of \citet{o:10.1371/journal.pone.0076911}, they are widely used.

@@

    \caption{Cumulative number of instances solved as a function of search space size, using four
    graph database datasets and a simple constraint programming subgraph isomorphism algorithm. Note
    that the $x$-axis shows recursive calls, and does \emph{not} use a log scale.}
    \label{figure:biiiig-data}
\end{figure}

Rather than the modern techniques used by Glasgow and LAD, we deliberately select a very simple
starting point: a constraint programming model implemented in Gecode 5.1.0 using only
toolkit-provided constraints and heuristics. The model uses a ``smallest domain first'' variable
ordering heuristic with tie-breaking on degree, an all-different constraint, extensional constraints
for adjacency, and labelled-degree filtering at the top of search. Even this simple approach finds
nearly every instance in each of the four datasets trivial, and no exponential behaviour is
observed. As we show in \cref{figure:biiiig-data}, the \emph{hardest} instance for the AIDS dataset
requires 287 recursive calls, and 149,268 of the instances can be solved without search.  For PDBS,
the hardest problem requires 3,161 recursive calls (which occurs 20 times, due to duplicated
queries). For both PCM and PPI, the hardest problem requires 23 recursive calls.  In other words,
none of these instances are in any way computationally hard even for a simple domain-based
algorithm, even before introducing stronger filtering, inference, or heuristics.  Furthermore it is
certainly not the case that a sequential scan is infeasible as so many filter~/ verify papers claim.
The exponential behaviour seen by others in these instances is purely down to a bad choice of
subgraph isomorphism algorithm, and these instances are not inherently ``really hard''.

>>>>

To show that a pure subgraph isomorphism approach is feasible, with no indexing or supporting
preprocessing, we look at four datasets commonly used in graph indexing evaluations. We do not claim
that these are high-quality datasets or that the associated queries are sensible, merely that
following the example of \citet{o:10.1371/journal.pone.0076911}, they are widely used. Each dataset
has labels on vertices (but not on edges), and we will look at the non-induced problem.

@@

    \caption{Cumulative number of instances solved as a function of search space size, using four
    graph database datasets, and either VF2 or a simple constraint programming subgraph isomorphism
    algorithm implemented in Gecode. The $x$-axis shows recursive calls, using a log scale, and each
    line ends at the $x$ value where every instance has been solved.}
    \label{figure:biiiig-data}
\end{figure}

Rather than the modern techniques used by Glasgow and LAD, we deliberately select a very simple
starting point: a constraint programming model implemented in Gecode 5.1.0 using only
toolkit-provided constraints and heuristics. The model uses a ``smallest domain first'' variable
ordering heuristic with tie-breaking on degree, an all-different constraint, extensional constraints
for adjacency, and labelled-degree filtering at the top of search. Even this simple approach finds
nearly every instance in each of the four datasets trivial, and no exponential behaviour is
observed. As we show in \cref{figure:biiiig-data}, the \emph{hardest} instance for the AIDS dataset
requires 287 recursive calls, and 149,268 of the instances can be solved without search.  For PDBS,
the hardest problem requires 3,161 recursive calls (which occurs 20 times, due to duplicated
queries). For both PCM and PPI, the hardest problem requires 23 recursive calls.  In other words,
none of these instances are in any way computationally hard even for a simple domain-based
algorithm, even before introducing stronger filtering, inference, or heuristics.  Furthermore it is
certainly not the case that a sequential scan is infeasible as so many filter~/ verify papers claim.
In contrast, with VF2, the hardest instance in PDBS requires 31,189 recursive calls; in AIDS,
168,569 calls; in PCMS, 6,476,072 calls; and in PPI, 2,913. In other words, any exponential
behaviour seen in these instances is purely down to a bad choice of subgraph isomorphism algorithm,
and these instances are not inherently ``really hard''.
